
\section{Discussion}

Axial vortices were created over a range of velocities, and interrogated at 
seven different stations ranging between 546$mm$ and 1016$mm$ downstream of the 
end of the vortex generators center body, representing distances between 5.4 
and 10 chord lengths. For the 
purposes of results and discussion, it is desirable to focus on a smaller group 
of exemplary data sets that illustrate observed phenomena. The vortex generator 
used in these experiments had a center body with a 1 inch diameter, and thus 
produced two co-rotating vortices that merge together as they progressed 
downstream. 
The first three stations display characteristics that resemble co-rotating 
vortices that had not yet finished merging. Furthermore, analysis shows that 
the uncertainty in measurements decreases significantly as the interrogation 
plane moves down stream. Therefore, results from only a few vortices at 
downstream locations are discussed here. Complete measurement sets for all 
vortices generated as part of this study can be obtained from an online public 
repository which can be accessed with its DOI,  
\url{http://dx.doi.org/10.5281/zenodo.46745}. Turbulent quantities have 
been compared with hot-wire anemometry experiments using a bodiless bi-wing 
vortex generator in the same wind tunnel \cite{thompson2016}.


\subsection{Vortex Stabilization}
A bi-wing vortex generator creates a pair of co-rotating vortices, emanating 
from the intersections of the wing and center body. A single 
axial vortex stabilizes as the co-rotating vortices merge together and as the 
wake of the center body diminishes. The rotational motion of the vortex 
winds the axial velocity deficit around the core region. This was observed 
in the PIV data in Figure \ref{fig:run_1_W_contour} showing a far upstream 
mean axial velocity distribution with very clear wing wake zones radiating from 
the center.Those rays largely disappear further downstream as shown in Figure 
\ref{fig:run_55_W_contour} at a higher velocity with wake remnants wound about 
the core. 

\vspace{32pt}
\input{figs/run_1/run_1_W_contour}
\input{figs/run_55/run_55_W_contour}

Each wing generated a lifting force which converted axial momentum into radial 
and tangential momentum about the vortex core. As the vortex developed, mean 
radial fluid motion persisted with similar magnitude, though began to break up 
into smaller localized jets as shown in Figures \ref{fig:run_1_R_contour} and 
\ref{fig:run_51_R_contour}, which were taken at similar free stream velocities 
but several chord lengths apart. For radial velocity components, positive 
values indicate fluid flowing away from the core. For the majority of tests, 
the radial velocity patterns from the wings remain intact, with radial outflow 
in the top right and bottom left quadrants. 

\vspace{32pt}
\input{figs/run_1/run_1_R_contour}
\input{figs/run_51/run_51_R_contour}

Under the conditions of early vortex formation, stronger asymmetry was seen in 
the azimuthal velocity component, $\overline v_{t}$ than in more developed 
vortices as 
shown in Figures \ref{fig:run_1_T_contour} and \ref{fig:run_55_T_contour}. When 
this data is mapped into a scatter plot for every grid point making up the 
two dimensional dataset, the asymmetries become particularly apparent as shown 
in Figure \ref{fig:run_1_T_vs_r_mesh_scatter}.

\vspace{32pt}
\input{figs/run_1/run_1_T_contour}
\input{figs/run_55/run_55_T_contour}
\input{figs/run_1/run_1_T_vs_r_mesh_scatter}

Vortices are characterized principally by the maximum azimuthal velocity 
and the associated radius of the core. The absolute maximum velocity from the 
scatter plot is likely to be impacted by random noise in the data, and may 
not occur near the true core boundary. Therefore, the vortices were 
characterized by sorting the scattered set of points of azimuthal velocity vs 
radius as shown in \ref{fig:run_1_T_vs_r_mesh_scatter} in ascending order and 
performing a Gaussian smoothing to create a smoothed profile as shown in Figure 
\ref{fig:run_55_comparison}. This Gaussian smoothing creates the average value 
for a point from the scattered data by using a weighted average that prefers 
nearby points, but will accept many points further away, allowing a smooth 
profile to be drawn across the range of point densities inherent in the 
dataset. The 
maximum azimuthal velocity and core radius was then obtained from the smoothed 
curve. Thus,
developed vortex planes with pronounced radial symmetry were better suited for 
further study. With a smoothed azimuthal velocity profile, the experimental 
data were compared to the theoretical profile models of common vortices 
introduced previously, as shown in Figure \ref{fig:run_55_comparison}. In this 
comparison, the Lamb-Oseen and Ash vortices are scaled to match the 
experimental maximum azimuthal velocity, while the Rankine vortex is scaled to 
match the core radius.

\vspace{32pt}
\input{figs/run_55/run_55_comparison}

The measured vortex characteristics are shown in Tables
\ref{table:experiment_results_1} through \ref{table:experiment_results_7} where 
$I_Z$ is the downstream distance, 
$V_{nom}$ is the nominal free stream velocity, $dt$ is the time step between 
image pairs, $V_{free}$ is the actual free stream velocity, $P_{atm}$ is the 
atmospheric pressure at the time of the test, 
$T_{tunnel}$ is the temperature measured inside the wind tunnel, $\phi$ is the 
relative humidity, $\eta_P$ is the pressure relaxation coefficient, $R_{core}$ 
is the core radius, $\overline{v_{\theta}}_{max}$ 
is the maximum azimuthal velocity, and $\overline{v}_{z}$ is the axial 
velocity averaged over the entire observation area.

\vspace{32pt}
\input{tables/experiment_results_1}
\input{tables/experiment_results_2}
\input{tables/experiment_results_3}
\input{tables/experiment_results_4}
\input{tables/experiment_results_5}
\input{tables/experiment_results_6}
\input{tables/experiment_results_7}

\subsection{Macro-Turbulent Structures}
\label{section:macroturb}

The vortex cores sizes at all tested velocities were observed to shrink from 
a radius of 17-20$mm$ 684$mm$ down stream to a radius of 13-16$mm$ 914$mm$ down 
stream, a span of just 50$mm$. The abrupt contraction of the core 
may be attributable to the instability created by the wake of the vortex 
generator center body. As the core diameter decreased, the azimuthal velocities 
were also observed to increase sharply as the fast moving fluid accelerates 
inward to tighten the core boundary. Figure \ref{fig:run_35_R_contour} shows 
nearly opposed and well-defined regions of radial velocity into the core (-) 
and out of the core (+). Figure \ref{fig:run_45_R_contour} shows the dispersion 
of these radial velocity jets just one half chord length further downstream, at 
the same free stream velocity. 

\vspace{32pt}
\input{figs/run_35/run_35_R_contour}
\input{figs/run_45/run_45_R_contour}

At this transition stage, an abrupt change in the turbulent 
kinetic energy distribution was also observed. Figure 
\ref{fig:run_35_ctke_contour} (at station 4) shows turbulent energy spread 
widely across a region approximately twice the diameter of the core, with 
several areas showing turbulent kinetic energy intensities of 95\% of the peak 
value in the core. Figure \ref{fig:run_45_ctke_contour} at station 5 shows a 
significant diminution of these zones of high intensity turbulence just outside 
the core region, while also indicating a 30\% increase in turbulent intensity 
within the core. An appearance of fresh turbulent bands spiraling radially
outwards was also observed. Perhaps even more interestingly, small areas of 
high turbulent intensity were found to arise in the same region as before, 
between 1 and 2 core radii, just another half chord length down stream as shown 
in Figure \ref{fig:run_55_ctke_contour} at station 6. In this figure, a drop 
in the maximum turbulence intensity was also observed in the center. This could 
indicate incomplete coalescence of a single vortex. Despite 
the new strengthening of turbulent structures outside the core region, the core 
radius was not observed to increase between stations 5 and 6. These findings 
are consistent with the vortex gulping phenomenon described by 
\cite{bandyopadhyay1991}. Furthermore, this suggests the ingestion of 
turbulence may be responsible for tightening the core, contributing to the 
longevity of axial wake vortices observed in nature.

\input{figs/run_35/run_35_ctke_contour}
\input{figs/run_45/run_45_ctke_contour}
\input{figs/run_55/run_55_ctke_contour}

The previous discussion is based upon data produced from time averages of many 
PIV frames at one second intervals. To gain better understanding of the 
relevant time and 
length scales at work in the turbulent regions of the flow, particularly inside 
the core and in the highly turbulent zones just outside the core, the unsteady 
(unstable) components were examined. The time series is comprised of just 200 
seconds of samples at 1Hz, enough to resolve motions at frequencies between 
about 0.01 Hz and 0.5 Hz. Understanding the characteristic time and length 
scales of turbulent fluctuations within the core, and how they relate to those 
seen outside of the core is of great interest. Evolution of turbulent scales as 
the interrogation plane moves downstream is of further interest. 

%ensures every page has "Figure         Page" at the top of it
\makeatletter
\addtocontents{lof}{%
	\protect\afterpage{\protect\hbox to \linewidth%
		{\noindent Figure \hfill Page}\par%
		\protect\vspace{12\p@}}}
\makeatother

Figures \ref{fig:run_35_ctke_01rdynamic} through 
\ref{fig:run_55_ctke_01rdynamic} show power spectral densities of 
average turbulent kinetic energies within the core volume for a vortex at the 
same three conditions as displayed in Figures \ref{fig:run_35_ctke_contour} 
through \ref{fig:run_55_ctke_contour}. Each of these figures contains three 
plots. The left-most plot shows the time averaged kinetic energy distribution 
within the sample area. The center plot shows the variation in kinetic energy 
over time with both the mean value of all points within the sample area shown, 
and 5th and 95th percentile boundaries to show the spread. The plot at right 
shows the log of the power spectral density (PSD) of the mean line shown in the 
center plot. 

\input{figs/run_35/run_35_ctke_01rdynamic}
\input{figs/run_45/run_45_ctke_01rdynamic}
\input{figs/run_55/run_55_ctke_01rdynamic}

With the region inside the core boundary highlighted, the pronounced 
concentration of turbulent energy within the core is clearly visible. Figures 
\ref{fig:run_35_ctke_12rdynamic} through \ref{fig:run_55_ctke_12rdynamic} show 
the magnitude of turbulent energy in the area just outside the core, between 
one and two core radii, to sharply decrease between 8.5 and 9.0 chord lengths 
down stream, and then remain the same. The 
power spectral density profile of the core region however did not appear to 
change significantly, with most frequencies remaining between -5 and -10 
$db/Hz$ at all three chord lengths. The power spectral density profiles were 
examined for the entire range of velocities at each station, and no distinct 
trends were discovered.

\input{figs/run_35/run_35_ctke_12rdynamic}
\input{figs/run_45/run_45_ctke_12rdynamic}
\input{figs/run_55/run_55_ctke_12rdynamic}

Another quantity of interest is the average turbulence within specific regions 
around the core. Average turbulent energy within the core was found to increase 
proportionally with free stream velocity for a constant downstream distance. 
The power spectral density reflected 
a similar trend of an even increasing magnitude across the frequency range. The 
distribution within the core however undergoes seemingly cyclical periods as 
velocity increases; between phases with high intensity turbulence spread in 
spokes extending to near the edges of the core boundary, and with the 
turbulence concentrated more tightly within half of the total radii as shown in 
Figures \ref{fig:run_51_ctke_01rdynamic} through 
\ref{fig:run_60_ctke_01rdynamic} at $z/c = 9.5$.

\input{figs/run_51/run_51_ctke_01rdynamic}
\input{figs/run_52/run_52_ctke_01rdynamic}
\input{figs/run_53/run_53_ctke_01rdynamic}
\input{figs/run_54/run_54_ctke_01rdynamic}
\input{figs/run_55/run_55_ctke_01rdynamic}
\input{figs/run_56/run_56_ctke_01rdynamic}
\input{figs/run_57/run_57_ctke_01rdynamic}
\input{figs/run_58/run_58_ctke_01rdynamic}
\input{figs/run_59/run_59_ctke_01rdynamic}
\input{figs/run_60/run_60_ctke_01rdynamic}

%ensures every page has "Figure         Page" at the top of it
\makeatletter
\addtocontents{lof}{%
	\protect\afterpage{\protect\hbox to \linewidth%
		{\noindent Figure \hfill Page}\par%
		\protect\vspace{12\p@}}}
\makeatother

A regular upward trend in peak turbulence at the cores center was observed with 
increasing free stream velocity. These were compared to hot-wire anemometry 
measurements at greater free stream velocities and up to 
14 chord lengths downstream, and found to be consistent \cite{thompson2016}.

Over the range of velocities tested, the time between frame captures, $dt$ was 
changed between nominal free stream velocities of 25 and 27 $m/s$. A smaller 
$dt$ value was required at higher velocities to ensure that some particles 
within the light sheet for the first image capture would remain there for the 
second. An uncertainty analysis demonstrated that, since the average particle 
moves fewer pixels in the shorter period of time, the precision in the 
displacement computations is reduced for a shorter $dt$. Since the Reynolds 
averaged fluctuating components used to calculate turbulent energy are directly 
related to the precision in the measurement, this change in $dt$ could have 
impacted the relationship between free stream velocity and turbulent kinetic 
energy.

The PIV measurements have an uncertainty associated with the mean components 
which benefits from an increased number of samples. However, the 
uncertainty associated with each individual measurement (N=1) has a 
significant impact on power spectral density analysis of the fluctuating 
values. An uncertainty analysis has been conducted and indicates that the 
precision of PIV measurements is sufficient to study time averaged quantities 
including Reynolds stresses and those related to the root-mean-square of 
velocity fluctuations, but insufficient to study the real time dynamic nature 
of turbulent structures. The basis for this conclusion is discussed in greater 
detail in Section \ref{sec:piv_uncert}.

\section{Reynolds stress and turbulence relationships}

Reynolds stresses represent components of the overall stress tensor, and have 
been used extensively in the study of turbulence. Turbulent viscosity models, 
which use a scalar viscosity term to represent the transport and dissipation of 
energy in small scale flows make direct assumptions about the Reynolds stresses 
in a given fluid flow. 

\clearpage
\subsection{Velocity fluctuations}

While turbulent energy and Reynolds stress, especially the radial-tangential 
reynolds stress, were very closely related, each component exhibited unique 
characteristics. Repeated patterns were observed in the distribution of 
velocity fluctuation intensities and turbulent energy. Since the sampling rate 
of the PIV system was 1Hz, an alternate approach was taken to understand the 
time scales of periodic structures. If turbulent structures are assumed to 
convect downstream, then profiles from different axial vortices can be compared 
on the basis of 'age'. In principle, a small increase in free stream velocity 
is analogous to taking a small step upstream to where the vortex is 'younger'. 

The following progression is shown one velocity step at 
a time in Figures \ref{fig:run_41_rr_contour} through 
\ref{fig:run_50_rr_contour}. In these figures, the color values are scaled to 
range from the minimum value to the 99 percentile value. This was done to 
preserve the dynamic range of coloring in the moderate intensity regions where 
the structures of greatest interest reside.
At 9 chord lengths down stream and 15 $m/s$, four lobes of 
large magnitude radial fluctuations could be seen faintly. At 17 $m/s$, the 
lobes appeared to turn in on each other to form two crescents radiating from 
the left and right sides of the vortex core. At 19 $m/s$, the C shapes were 
observed to split, giving the plot an open four leaf clover shape. At 21 $m/s$, 
the four leaves appear to finish closing on themselves to form four 
elliptical lobes extending about one core radii. At 23 $m/s$, these 
structures appeared to break up, only for familiar C shapes to arose again at 
25, 27, and 29 $m/s$. By 31 and 33 $m/s$ the C shaped protrusions closed and 
became elliptical in appearance once more.  The size of the 
elliptical lobes between 21 $m/s$ and 33 $m/s$ were not distinguishably 
different. 

\input{figs/run_41/run_41_rr_contour}
\input{figs/run_42/run_42_rr_contour}
\input{figs/run_43/run_43_rr_contour}
\input{figs/run_44/run_44_rr_contour}
\input{figs/run_45/run_45_rr_contour}
\input{figs/run_46/run_46_rr_contour}
\input{figs/run_47/run_47_rr_contour}
\input{figs/run_48/run_48_rr_contour}
\input{figs/run_49/run_49_rr_contour}
\input{figs/run_50/run_50_rr_contour}

With this perspective, a consistent sizing of periodic structures in velocity 
fluctuations is indicative of a gradual evolution. If it can be assumed that 
vortex cross sections of the same 'age' are similar, the frequency of the 
periodicity indicated by this repeated pattern can be found by

\begin{equation}
f = \Big(\frac{Z_1}{V_1} - \frac{Z_1}{V_2}\Big)^{-1} = 
\Big(\frac{0.9144m}{21.18 m/s}-\frac{0.9144m}{33.05 m/s}\Big)^{-1} \approx 63 Hz
\label{eq:period_frequency}
\end{equation}

\noindent
Where $V_1$ is the lower velocity, $V_2$ is the greater, and $Z$ is the 
position of the interrogation plane. A similar periodicity of 57Hz was easily
observed in radial fluctuations at station 6. The PIV system used for these 
experiments sampled at 1Hz, well below the sampling rate required to observe 
this periodicity by examination of power spectral density of the velocity 
fluctuations. The radial component of the velocity fluctuations in the core 
region are strongly influenced by meandering of the cores center axis 
\cite{bandyopadhyay1991}. The 
high intensity region within 1/2 core radius could be largely due to wandering, 
while regions well outside the core region will be influenced very slightly by 
wandering.

Similar patterns such as the set of four protruding elliptical structures were 
observed in the axial velocity fluctuations as shown in 
Figure \ref{fig:run_50_ww_contour}, but were not as pronounced as those in the 
radial direction and did not demonstrate periodicity with the same level of 
clarity.

\input{figs/run_50/run_50_ww_contour}

Interestingly, the azimuthal velocity fluctuations demonstrated some
periodic fluctuations as well as some very consistent zones of low fluctuation 
intensity.
Figures \ref{fig:run_41_tt_contour} through \ref{fig:run_50_tt_contour} 
consistently show areas of very low tangential velocity fluctuation on the left 
and right sides of the vortex core, each between 1 and 2 times the size of the 
vortex core. A cyclical appearance and disappearance of high intensity 
fluctuations occurs in the large regions above and below the core with the 
vector fields at 19, 23, and 29 $m/s$ exhibiting very little azimuthal velocity 
fluctuations in the regions above and below.

\input{figs/run_41/run_41_tt_contour}
\input{figs/run_42/run_42_tt_contour}
\input{figs/run_43/run_43_tt_contour}
\input{figs/run_44/run_44_tt_contour}
\input{figs/run_45/run_45_tt_contour}
\input{figs/run_46/run_46_tt_contour}
\input{figs/run_47/run_47_tt_contour}
\input{figs/run_48/run_48_tt_contour}
\input{figs/run_49/run_49_tt_contour}
\input{figs/run_50/run_50_tt_contour}

%ensures every page has "Figure         Page" at the top of it
\makeatletter
\addtocontents{lof}{%
	\protect\afterpage{\protect\hbox to \linewidth%
		{\noindent Figure \hfill Page}\par%
		\protect\vspace{12\p@}}}
\makeatother


Using a similar vortex age approach as Equation \ref{eq:period_frequency}, the 
frequency of this motion was estimated to be 120Hz, approximately twice that of 
the observed frequency of periodic radial fluctuations. Each of these 
periodicities is more clearly visible than that observed in turbulent energies 
within the core as described in Section \ref{section:macroturb}, but are likely 
related. The wind tunnel flow is driven by a direct current electric motor, and 
the speed is controlled by voltage. No evidence has been found to suggest these 
findings have been influenced by the utility frequency, which is also 60Hz.

\subsection{Constant turbulent viscosity}
The turbulent viscosity hypothesis first introduced by 
Boussinesq \cite{pope2000} in the notation of Pope, can be described by

\begin{equation}
-\rho \langle u_i u_j \rangle + \frac{2}{3}\rho k \delta_{ij} = \rho \nu_T
\frac{\partial \langle U_i \rangle}{\partial x_j} + \frac{\partial \langle U_j 
\rangle}{\partial x_i}
\end{equation} 

\noindent
which assumes the deviatoric Reynolds stress is proportional to the mean rate 
of strain. Because the non-equilibrium vortex model proposed by Ash, Zuckerwar, 
and Zardkahan in \cite{ash2011} is an exact solution the Reynolds averaged 
Navier-Stokes equations, and very accurately predicts the azimuthal velocity 
profile of a turbulent axial wake vortex, it was of interest to examine the 
validity of 
the turbulent viscosity hypothesis with experimental data. All terms in the 
summation of this equation were calculated from the PIV dataset, with the 
exception of gradients in the axial, $Z$, direction. Neglecting those terms, 
the turbulent viscosity assumption was examined for each vortex dataset. The 
turbulent viscosity calculation involves division by spatial derivatives, which 
can locally be very close to zero. In the area within 1.5 core radii where the 
velocity gradients were consistently high, the turbulent viscosity calculation 
was subject to much less noise than regions far from the core with small 
gradients as shown in figure \ref{fig:run_55_turb_visc_vs_r_meshscatter}. 
Calculations were performed in Cartesian coordinate space, and the values 
of derivatives near the $X$ and $Y$ axes through the center of the vortex 
tended to be near zero, and caused unrealistically high values to occur very 
frequently in these zones with no physical significance. To reduce noise, the 
scatter plots shown do not include points within 5 degrees of a vertical or 
horizontal axis.

\vspace{32pt}
\input{figs/run_55/run_55_turb_visc_vs_r_meshscatter}

\noindent
Within the core region, a one or two equation turbulence model where $\nu_T$ is 
a function of the velocity gradients, such as the $k-\epsilon$ model is likely 
required. While the scatter plot is very noisy outside the core region, 
90\% of values consistently fell within a band of small positive values between 
0 and 0.1 $m^2/s$, which is 5 orders of magnitude larger than expected.
These trends were observed consistently in all vortices.

\subsection{Non-equilibrium turbulence}
The solution describing the behavior of an axial wake vortex derived from 
non-equilibrium pressure theory also presents a definition for turbulent 
viscosity. The definition can be derived from the Reynolds stress form of the 
cylindrical conservation of linear momentum equations and the continuity 
equation. In the individual Reynolds component notation used previously, 
turbulent viscosity derived from non-equilibrium pressure theory for an axial 
vortex can be expressed as

\begin{equation}
\nu_T =
    \cfrac{\cfrac{1}{r^2}\cfrac{d}{dr}\Big[r^2 (\overline{v_{\theta}^\prime 
    v_{r}^\prime})\Big] +
	\cfrac{\eta_p \bar{v_\theta}}{r^2}\Bigg[\overline{v_{r}^\prime 
	v_{r}^\prime} - 
	\overline{v_{\theta}^\prime v_{\theta}^\prime} + 
	r \cfrac{d(\overline{v_{r}^\prime 
	v_{r}^\prime})}{dr}\Bigg]} {\Bigg[\cfrac{d^2 \bar{v_{\theta}}}{dr^2} + 
	\cfrac{d}{dr}
	\Bigg(\cfrac{\bar{v_{\theta}}}{r}\Bigg)\Bigg]}
\label{eq:etap_nut}
\end{equation}

\noindent
The magnitude of the pressure relaxation term is expected to be small compared 
with the first term with the radial derivative of the azimuthal-radial Reynolds 
stress, due in part to the typical microsecond order of the pressure relaxation 
coefficient, $\eta_p$. All radial gradient terms are expected to be quite 
large, especially in the vicinity of the core. The relative magnitudes of the 
Reynolds stress term in the numerator and the azimuthal velocity gradient terms 
in the denominator required investigation. Each of the three 
terms was examined individually. The first Reynolds stress term in the 
numerator was found to increase at a greater than logarithmic rate as 
$R/r_{core}$ decreased as shown in Figure 
\ref{fig:run_55_turb_visc_reynolds_vs_r_meshscatter}. As expected, the 
magnitude of the pressure relaxation term was consistently found to be many 
orders smaller as shown in Figure 
\ref{fig:run_55_turb_visc_ettap_vs_r_meshscatter}.

\vspace{32pt}
\input{figs/run_55/run_55_turb_visc_reynolds_vs_r_meshscatter}
\input{figs/run_55/run_55_turb_visc_ettap_vs_r_meshscatter}

The term in the denominator related to the mean strain rate, a profile similar 
to the Reynolds stress term was observed, 
though the slope was found to be slightly shallower as shown in Figure 
\ref{fig:run_55_turb_visc_vel_grad_vs_r_meshscatter}.

\vspace{32pt}
\input{figs/run_55/run_55_turb_visc_vel_grad_vs_r_meshscatter}


Since the dominant terms in the numerator and denominator are of similar orders 
of magnitude, the resulting turbulent viscosity loses the strong logarithmic 
nature, but is subject to significant noise. The experimentally 
calculated turbulent viscosity utilizing all the terms as in Equation 
\ref{eq:etap_nut} was found too scattered to be of use, particularly in the 
vicinity of the vortex core as shown in Figure 
\ref{fig:run_55_turb_visc_total_vs_r_meshscatter}. Spatial derivatives in 
cylinrdical coordinates were obtained by a chain rule product between a 
Cartesian spatial derivative and the sin or cosine of $\theta$, leaving many 
opportunities for velocities at small angles or small errors in denominators to 
become very large errors. The opportunities for these errors to arise becomes 
greater in the vicinity of the core.

\vspace{32pt}
\input{figs/run_55/run_55_turb_visc_total_vs_r_meshscatter}

This exercise was useful in showing the relative magnitudes of the strain rate 
and Reynolds stress terms in Equation \ref{eq:etap_nut}, but did not yield an 
explanation for why a constant turbulent viscosity assumption can be used to 
accurately approximate the velocity profiles of an axial wave vortex.

More simply, a total viscosity can be estimated by 
substituting circulation strength Reynolds number, $R_\Gamma$, and rearranging 
Equation \ref{eq:ash_vortex_model} for total turbulence, $\nu$, by

\begin{equation}
\nu(r) = \Bigg[\frac{v_\theta(r) \sqrt{\eta_P}}{2 \sqrt{2}}
\frac{(r/r_{core})^2 + 1}{(r/r_{core})}\Bigg]^2
\label{eq:total_visc}
\end{equation}

Viscosity was calculated from the mean azimuthal velocity, and exhibited an 
evolving behavior for varied free stream velocities. At low free stream 
velocities, total viscosity remained fairly constant near the kinematic 
viscosity of air, with slight increases typically observable between 1 and 1.5 
core radii and significant increases at less than 0.5 core radii as shown in 
Figures 
\ref{fig:run_51_turb_visc_by_vtheta_vs_r_meshscatter} through 
\ref{fig:run_55_turb_visc_by_vtheta_vs_r_meshscatter}. At a free stream 
velocity of 27 m/s, two small bumps in total viscosity are present near 1 and 2 
core radii respectively, and the profile becomes very scattered beyond 2.5 core 
radii as shown in Figure \ref{fig:run_56_turb_visc_by_vtheta_vs_r_meshscatter}, 
possibly indicating an incomplete vortex merger. A more extreme scattering and 
divergence at large core radii continued to be seen with further increasing 
free stream velocity. Tests at lower free stream velocities consistently showed 
that total viscosity, and thus turbulent viscosity could be well approximated 
by a constant value. 

\vspace{32pt}
\input{figs/run_51/run_51_turb_visc_by_vtheta_vs_r_meshscatter}
\input{figs/run_52/run_52_turb_visc_by_vtheta_vs_r_meshscatter}
\input{figs/run_53/run_53_turb_visc_by_vtheta_vs_r_meshscatter}
\input{figs/run_54/run_54_turb_visc_by_vtheta_vs_r_meshscatter}
\input{figs/run_55/run_55_turb_visc_by_vtheta_vs_r_meshscatter}
\input{figs/run_56/run_56_turb_visc_by_vtheta_vs_r_meshscatter}
\input{figs/run_57/run_57_turb_visc_by_vtheta_vs_r_meshscatter}
\input{figs/run_58/run_58_turb_visc_by_vtheta_vs_r_meshscatter}
\input{figs/run_59/run_59_turb_visc_by_vtheta_vs_r_meshscatter}
\input{figs/run_60/run_60_turb_visc_by_vtheta_vs_r_meshscatter}

While 
total viscosity generally increased with free stream velocity, the relationship 
was not as consistent as expected. Equation \ref{eq:total_visc} shows that 
$\nu$ is related linearly to the pressure relaxation coefficient, 
$\eta_P$, which is very strongly dependent upon relative humidity 
\cite{ash2011}. Relative humidity was calculated from wet bulb and dry bulb 
measurements were taken with a sling psychrometer, and recorded with a 
precision no greater than 0.5 degrees Fahrenheit. Variations in relative 
humidity are likely to have caused sufficient uncertainty in the pressure 
relaxation coefficient to cause shifts in the total viscosity calculation. A 
simple comparison revealed that the viscosity seen at low free stream 
velocities was significantly less than the kinematic viscosity of air, which is 
very unlikely to be true. Therefore, kinematic 
viscosity was not subtracted from the total viscosity to yield a measurement of 
the effective turbulent viscosity.
\pagebreak