\section{PIV Overview}The present study uses a stereo PIV system which resolves three dimensional near-instantaneous velocity vectors gridded on a two dimensional cross section of flow. The cameras are capable of taking two images just a few microseconds apart. Determining velocity vectors requires two images just a few microseconds apart, but pairs of images can be taken at greater time intervals. The PIV method used in this study utilizes a "frame-straddling" technique that spaces the time interval between two laser pulses so that the camera sensor arrays receive the first laser illuminated image data as near as possible to the closing of the shutter, followed by a second laser illuminated image as close as possible to the beginning of the second shutter opening. This technique allows both cameras to capture image sequence pairs simultaneously, spaced 25 to 50 microseconds apart. This can be repeated once every second, resulting in a true velocity sampling frequency of 1 Hz.\subsection{Seeding the Flow}Appropriate particle seeding density and time between straddled frames is thesubject of continued study, and is difficult to predict \textit{a priori}. Complete coverage of a two dimensional vector field is highly dependent upon uniform optimal particle density conditions which are difficult to obtain, and maintain over an extended test interval, due to seed-particle accumulation. For stereo PIV, incomplete data in either of the two dimensional vector sets from either camera at a given spatial location will result in an indeterminate vector displacement in the three dimensional vector data. To elevate the likelihood that a displacement vector at a given location can be properly determined, an additional data refining technique outlined by \cite{hart1998} was employed. The Hart method compares correlation maps between adjacent vector spots to produce fewer errors than would otherwise be produced by completely independent evaluation of each small image sector. In instances where two adjacent regions lack a well-defined peak, the Hart method emphasizes shared peaks in order to reveal a significant correlation that might otherwise have been missed. In instances where sub optimal seeding conditions exist and a correlation map produces a false peak, the Hart method can isolate and eliminate those anomalous peaks based on comparison with adjacent sectors \cite{hart1998}. The actual process by which the Hart method reduces erroneous vectors is difficult to quantify on a case by case basis, but is expected to have a positive effect and reduce overall uncertainty in the PIV measurement.