\tolerance=10000

% general packages that seem to be required by everything
\documentclass[12pt]{report}
\usepackage{geometry}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}

\usepackage{ODUthesis} 		% custom template specific to ODU thesis.
\usepackage{indentfirst} 	% indents first line of a section (style guide)
\usepackage{hyperref} 		% allows use of hyperlinks
\usepackage{gensymb} 		% allows use of degree symbol with \degree
\usepackage{longtable} 		% formats tables
\usepackage{float} 			% allows use of figure{}[H]
\usepackage{subcaption} 	% allows mutliple figures for one caption
\usepackage{mathtools}		% mathtools including multi lined equations
%\usepackage{longtable}		% allows long tables that span multiple pages

\usepackage[font=small, labelfont=bf]{caption} 	% formats figure captions

\graphicspath{{figs/}} 		% defines reference path for images and figures

%This style follows conventions used in Physical Review C. The names of 
%figures, tables and captions can be changed by using the commands below with 
%appropriate changes. For example, "FIG." could be replaced by "Fig." or 
%"Figure" to change the labeling of figure captions.

%\renewcommand{\figurename}{FIG.}
%\renewcommand{\tablename}{TABLE}
%\renewcommand{\bibname}{BIBLIOGRAPHY}

\begin{document}

\title{Experimental Investigation of Turbulent Structures and Non-Equilibrium 
Effects in Axial Wake Vortices Via Particle Image Velocimetry}
\author{Jeffry William Ely}
\principaladviser{Robert Ash}
\member{Drew Landman} 
\member{Colin Britcher} 
\degrees{Bachelor of Science in Mechanical Engineering}
\dept{Aerospace Engineering}
\submitdate{May 2016}
\vita{Bachelors of Mechanical Engineering, Old Dominion University, December 
2011}

\phdfalse 			%produces language on title page for Masters Thesis
%\copyrightfalse    %suppresses copyright notice
%\figurespagefalse  %suppresses List of Figures
%\tablespagefalse   %suppresses List of Tables

%=========================================================================
\abstract{
	Vortices are a common phenomenon in fluid flows that arise as kinetic 
	energy dissipates into heat via viscous interaction. They arise naturally 
	at large scales in the form of dust devils, tornadoes, and in the wake of 
	aircraft. It is important to understand the conditions leading to their 
	formation, their duration, and their dissipation to prevent undesirable 
	effects. Among these effects is an decrease in safety of aircraft 
	operations in the wake of other aircraft, an extremely common situation at 
	airports around the world. A large number of mathematical models and 
	experimental data exists to help explain various aspects of axial wake 
	vortex behavior, but current models fail to understand why many vortices 
	remain tightly wound about their cores for as long as they have been 
	observed to do. The current study builds upon the theoretical work of Ash, 
	Zardadkhan and Zuckerwar \cite{ash2011}, and tests specific attributes of a 
	vortex for agreement with non-equilibrium pressure relaxation theory. This 
	theory provides an exact solution to the Navier-Stokes equations for an 
	axial vortex, with a resulting velocity model that agrees with leading 
	empirical models. A bi-wing vortex generator was constructed, and placed in 
	a low speed wind tunnel. Stereo particle image velocimetry was employed to 
	map three dimensional velocity vectors in a plan perpendicular to the free 
	stream flow direction at a rate of 1Hz for 200 seconds. A Reynolds time 
	averaging approach was used to synthesize this velocity data into stable 
	and fluctuating components at multiple Reynolds numbers and downstream 
	locations. 
	}

\beforepreface

%\prefacesection{Acknowledgements}

\prefacesection{Notes}
Good practice for preserving the code used to perform this research was 
attempted by use of git version control software and a git repository. All code 
used to perform data analysis and visualization, along with the raw three 
dimensional vector dataset can be downloaded from 
\url{https://github.com/Jwely/pivpr}.
	

\afterpreface

%=========================================================================
% introduction chapter and subdocuments
\chapter{Introduction}
Vortices are common natural fluid phenomenon which are extensively scrutinized 
in the study of turbulent flows. While small vortical structures appear and 
disappear without notice, some circumstances promote the formation of strong 
axial vortices that endure. Some such axial vortices are found trailing the 
wake of aircraft wingtips, and in nature as dust devils, tornadoes, and 
waterspouts. Wingtip vortices behind large aircraft create unsafe conditions 
for other aircraft, especially around airport runways. The presence of axial 
wake vortices is a significant limiting factor on the volume of traffic an 
airport can accommodate \cite{hallock1991}. While the circumstances that result 
in the creation of strong axial vortices are well understood, the mechanisms 
that allow them to stay tightly wound for as long as they do remain mysterious. 
Due to the spontaneous nature of tornadoes, good experimental data is difficult 
to acquire and repeat. Axial wake vortices behave differently for 
small aircraft or in wind tunnels at low Reynolds number than they do for large 
aircraft or at high Reynolds numbers \cite{burnam2013}. Multiple empirical and 
analytical exact solutions exist to describe the velocity profiles of axial 
vortices, each with varying levels of skill.

In this analysis, A bi-wing vortex generator was constructed, and placed in 
a low speed wind tunnel. Stereo particle image velocimetry was employed to 
map three dimensional velocity vectors in a plan perpendicular to the free 
stream flow direction at a rate of 1Hz for 200 seconds. A Reynolds time 
averaging approach was used to synthesize this velocity data into stable 
and fluctuating components. Free stream velocity and downstream distance 
were varied to create a vortices at a variety of Reynolds numbers and produce a 
three dimensional perspective. Observed velocity profiles were compared against 
the predictions of multiple common vortex models. Relationships between 
Reynolds stresses, turbulent energies, strain rates, and dissipation were
examined to gain understanding of the compact core behavior, and to help 
explain the models performance.

\input{docs/intro/vortices}
\input{docs/intro/piv_fundamentals}

%=========================================================================
% experiment setup chapter and sub documents
\chapter{Experiment Setup}
A complete particle image velocimetry system was installed in the ODU low speed 
wind tunnel and employed to measure three-dimensional velocity fields produced 
by an axial vortex at multiple nominal velocities $V_{nom}$ and multiple 
interrogation planes. The interrogation planes were defined by their 
distance downstream of trailing edges of the bi-wing the vortex generator 
$I_Z$. Nominal wind tunnel velocity was varied from 15$m/s$ to 33$m/s$, 
sampling 10 distinct velocities in increments of 2$m/s$. The interrogation 
plane was moved at irregular intervals from 546$mm$ to 1016$mm$ as shown in 
table \ref{table:test_matrix_table}. This chapter discuss details of the 
experimental setup used to produce these datasets, including wind tunnel 
control, vortex generator setup, PIV system calibration, data acquisition, data 
processing, and data quality control.

\input{tables/test_matrix_table}

\input{docs/experiment_setup/lswt}
\input{docs/experiment_setup/vortex_generator}
\input{docs/experiment_setup/piv_overview}
\input{docs/experiment_setup/piv_acquisition}
\input{docs/experiment_setup/piv_processing}
%=========================================================================

% experiment results and sub documents
\chapter{Experiment Results}

A large quantity of data was collected across many variant parameters. For the 
purposes of results and discussion, it is desirable to reduce the set to a few 
exemplary data sets that illustrate observed phenomena. The vortex generator 
used in these experiments had a center body with a 1 inch diameter, and thus 
produced two co-rotating vortices that merge together as they move downstream. 
The first three stations display characteristics that resemble co-rotating 
vortices that had not yet finished merging. Furthermore, analysis shows that 
the uncertainty in measurements increases significantly as the interrogation 
plane moves upstream. Data taken at the first three stations can thus be 
excluded from the discussion of results.

The uncertainty analysis also demonstrates that, while the time averaged 
velocity components are relatively precise and unbiased, the uncertainty in the 
measurement of any one of the 200 samples was very poor, typically on the order 
of $\pm 1.0 m/s$. The unexpectedly high uncertainty in individual measurements 
combined with the low sampling rate of 1Hz yielded a data set with limited 
utility for the examination of turbulence, but results will be compared with 
similar hot-wire anemometry experiments conducted by Thompson 
\cite{thompson2016}.

\input{docs/experiment_results/noneq_pressure}
\input{docs/experiment_results/piv_uncertainty}
\input{docs/experiment_results/discussion}
\input{docs/experiment_results/conclusions}


%=========================================================================
% constructs the bibliography
\bibliographystyle{apalike}
\bibliography{bib/vortices,bib/other,bib/particles,bib/piv}

\begin{thebibliography}{99}
%This command adds some extra space in the table of contents
\addtocontents{toc}{\vspace*{12pt}}  

% adds an entry for the bibliography in the table of contents
\addcontentsline{toc}{chapter}{BIBLIOGRAPHY}  
\end{thebibliography}
%=========================================================================

\appendix
includes here

%\include{docs/appendices/piv_data_appendices}	
%\include{docs/appendices/uncertainty_appendices}
% starts blank page
\newpage

% initiates printing of the vita page
\vitapage

\end{document}
