\thispagestyle{plain}
\begin{center}
    \Large
    \textbf{Experimental Investigation of Turbulent Structures and 
    Non-Equilibrium Effects in Axial Wake Vortices via Particle Image 
    Velocimetry}
    
    \vspace{0.4cm}
    \large
%    Thesis Subtitle
    
    \vspace{0.4cm}
    Jeffry William Ely
    
    \vspace{0.9cm}
    \textbf{Abstract}
\end{center}


Vortices are a common phenomenon in fluid flows that arise to dissipate kinetic energy into heat via viscous interaction. They arise naturally at large scales in the form of dust devils, tornadoes, and in the wake of aircraft. It is important to understand the conditions leading to their formation, their duration, and their dissipation to prevent undesirable effects. Among these effects is an decrease in safety of aircraft operations in the wake of other aircraft, an extremely common situation at airports around the world. A large number of mathematical models and experimental data exists to help explain various aspects of axial wake vortex behavior, but current models fail to understand why many vortices remain tightly wound about their cores for as long as they have been observed to do. The current study builds upon the theoretical work of Ash, Zardadkhan and Zuckerwar \cite{ash2011}, and tests specific attributes of a vortex for agreement with non-equilibrium pressure relaxation theory. A bi-wing vortex generator was constructed, and placed in a low speed wind tunnel. Stereo particle image velocimetry was employed to map three dimensional velocity vectors in a plan perpendicular to the free stream flow direction at a rate of 1Hz for 200 seconds. A Reynolds time averaging approach was used to synthesize this velocity data into stable and fluctuating components. Free stream velocity and downstream distance were varied to create a variety of vortices, and to ensure the availability of high quality data.
